---
# Documentation: https://sourcethemes.com/academic/docs/managing-content/

title: "Note10 Kaplan Meier Estimation"
subtitle: ""
summary: ""
authors: []
tags: ['bds722','bds722note','km-estimation']
categories: []
date: 2020-02-18T21:53:55-06:00
lastmod: 2020-02-18T21:53:55-06:00
featured: false
draft: false

# Featured image
# To use, add an image named `featured.jpg/png` to your page's folder.
# Focal points: Smart, Center, TopLeft, Top, TopRight, Left, Right, BottomLeft, Bottom, BottomRight.
image:
  caption: ""
  focal_point: ""
  preview_only: false

# Projects (optional).
#   Associate this post with one or more of your projects.
#   Simply enter your project's folder or file name without extension.
#   E.g. `projects = ["internal-project"]` references `content/project/deep-learning/index.md`.
#   Otherwise, set `projects = []`.
projects: []
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- KM Estimation is a non-parametric estimation of survival function.

- when no event times are censored, a non-parametric estimator of S(T) is $1-F_{n}(t)$, where $F_{n}(t)$ is the empirical cumulative distribution function.  

- when some observations are censored, we can estimate S(t) using KM estimation.

Let: $n_i$ be number of objects at risk of dying at time $t_i$  
$d_i$ be number of events that occur at time $t_i$  

## <span style="color:red"> KM estimator </span>  
$\hat{S}(t) = \prod_{k=1}^i \frac{n_k - d_k}{n_k}$


## <span style="color:red"> Basic tests for difference in survival curves

Want to test:  
$H_0: S_1(t) = S_2(t)$  
$H_1: S_1(t) \neq S_2(t)$

```{r}
mydat <- matrix(c('d1i','d2i','n1i-d1i','n2i-d2i'),ncol = 2, byrow = T)
colnames(mydat) <- c('Group 1', 'Group 2')
rownames(mydat) <- c('Died','Survived')
mydat
```


Test Statistic:  
$Q = \frac{(\sum w_i (d_{1i} - \hat{e}_{1i}))^2}{\sum w_i ^2 \hat{v}_{1i}}$
where:  
$\hat{e}_{1i} = \frac{n_{1i}d_{1i}}{n_{i}}$ 

$\hat{v}_{1i} = \frac{n_{1i}n_{2i}d_i (n_i - d_i)}{n_i ^2 (n_i - 1)}$  

$w_i =$ some weighting functing

Tests:  
Log-rank: $w_i = 1$  
Wilcoxon: $w_i = n_i$  
Tarone-ware: $w_i = \sqrt{n_i}$  
Peto-Peto: $w_i = \prod_{i=1}^T (\frac{1-\hat{e}_i}{n_i + 1})$